Gozde and Liora Replication Code

### 2015-04-27 replication project.Rmd

---
---
title: "Gozde and Liora Replication Code"
author: "Gozde Guran and Liora Goldensher"
date: "3/6/2015"
output: html_document
---

@@

Note: Please refer to saved files in our repository for images from the article. Also please refer to the table and figures (Table 1, Figure 1, 2) in the repository. Image scales are rearranged in saved versions. Our regression values are slightly different from the authors as his stata code used a slightly different method than our function (glmmPQL) to calculate Penalized Quasi-likehood. Finally, we were not able to reproduce the last two figures that showed the joint impact of more than 2 variables; we'll keep working on these.

@@

# Figure 1 (Conditonal Effects of Unemployment and Immigration)
![Figure 1](figures/fig1.png)
```{r}
library(msm)
asy <- seq(from = min(data$z_asylumseeker), to = max(data$z_asylumseeker), by = .01)) # vector with possible values for asylumseekers

@@

# Figure 2 (Conditonal Effects of Salience and Variance)
![Figure 2](figures/fig2.png)
```{r}
# Graph 1
salience <- seq(from = min(data$salience_mean_c), to = max(data$salience_mean_c), by=.1) # vector with possible values for salience
variable <- salience
slopes <- model$coef$fixed["var_c"] + model$coef$fixed["salience_mean_c:var_c"] * variable # Estimated slopes based on estimated coefficients of salience (21) and salience x variance (39)
estmean <- model$coef$fixed # Estimated means for variables 
estvar <- vcov(model) # Estimated covariances
se <- rep(NA, length(variable)) # Creating empty vector for predicted s.e. values
# Creating a loop with deltamethod to calculate estimated standard errors
for (i in 1:length(variable)) {
  j <- variable[i]
  se[i] <- deltamethod (~ (x21) + (x39)*j, estmean, estvar)
}
# Create confidence intervals and combine all values in matrix.
upper <- slopes + 1.96*se
lower <- slopes - 1.96*se
variable.data <- cbind(variable, slopes, upper, lower)

@@

#Extension
Our extension was inspired by Andreas Wimmer and Yuval Feinstein's 2010 paper "The Rise of the Nation-State Across the World, 1816-2001."  This paper seperately tests the prevailing explanatory theories of the rise of the nation-state as the dominant form of governance.  Our extension takes a similar approach, testing the major theories of support for extreme right parties in the Western European context in individual models.  We operationalize the variables in these theories using the available measures from the original data set, and each model retains the measures of salience (the prominence of extreme right issues on parties' agendas) and variance (degree of consensus within parties) and their interaction as independant variables.

# Model 1: Adorno Thesis (Personality Traits and Value Orientations)
The first model is based on Theodor Adorno's 1950 theory that support for extreme right parties (in his case, fascist parties) originates from supporters' personality traits and value orientations, which make individuals more receptive to the appeals of extreme right parties.  We operationalize "personality traits and value orientations" using the individual level characteristics from our data set.

```{r}
model1 <- glm(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + partner + left_right_scale + max_er + salience_mean_c + var_c + var_c:salience_mean_c + country - 1)

@@

# Model 2: Parsons Thesis (Social Disintegration)
Parsons' 1942 theory of social disintegratoin posits that anomia, a breakdown of social norms, and attendant feelings of anxiety, anger and isolation brings about a longing for strong leadership and rigid ideologies that extreme right parties can provide.  While we do not have measures for the emotions of participants in the studies whose data we use, in our second model we examine the relationship between extreme right voting and the kinds of sentiments about social change (satisfaction with democracy, opinions about the European Union) that Parsons refers to.

```{r}
model2 <- glm(ext_vote ~ eu_neg + z_dem_satis + max_er + salience_mean_c + var_c + var_c:salience_mean_c + country - 1)
```

# Model 3: Group Conflict Model (Various theories)
Group conflict theories encompass a variety of theories.  Some emphasize the role of instrumental rationality, arguing that xenophobia results from a contest between immigrants and residents over scarce resources like jobs and welfare benefits. Others emphasize the scapegoating of immigrants as perceived threats to residents' status, power or resources.  In this last model, we use relevant variables representing both the availability of resources like jobs and welfare benefits and the prevalence of immigration.

```{r}
model3 <- glm(ext_vote ~ dispro_elect + federalism + z_asylumseeker + z_unemp + z_asylumseeker:z_unemp + z_replacement + z_replacement:z_unemp + z_replacement:z_asylumseeker + max_er + salience_mean_c + var_c + var_c:salience_mean_c + country - 1)

>>>>

---
title: "Gozde and Liora Replication Code"
author: "Gozde Guran and Liora Goldensher"
date: "3/6/2015"
output: html_document
---

@@

Note: Please refer to saved files in our repository for images from the article. Also please refer to the table and figures (Table 1, Figure 1, 2) in the repository. Image scales are rearranged in saved versions. Our regression values are slightly different from the authors as his stata code used a slightly different method than our function (glmmPQL) to calculate Penalized Quasi-likehood. Finally, we were not able to reproduce the last two figures that showed the joint impact of more than 2 variables.

@@

# Figure 1 (Conditonal Effects of Unemployment and Immigration)
![Figure 1](figures/fig1.png)
```{r}
library(msm)
asy <- seq(from = min(data$z_asylumseeker), to = max(data$z_asylumseeker), by = .01) # vector with possible values for asylumseekers

@@

# Figure 2 (Conditonal Effects of Salience and Variance)
![Figure 2](figures/fig2.png)

# Figure 2, graph 1
salience <- seq(from = min(data$salience_mean_c), to = max(data$salience_mean_c), by=.1) # vector with possible values for salience
variable <- salience
slopes <- model$coef$fixed["var_c"] + model$coef$fixed["salience_mean_c:var_c"] * variable # Estimated slopes based on estimated coefficients of salience (21) and salience x variance (39)
estmean <- model$coef$fixed # Estimated means for variables 
estvar <- vcov(model) # Estimated covariances
se <- rep(NA, length(variable)) # Creating empty vector for predicted s.e. values
# Creating a loop with deltamethod to calculate estimated standard errors
for (i in 1:length(variable)) {
  j <- variable[i]
  se[i] <- deltamethod (~ (x21) + (x39)*j, estmean, estvar)
}
# Create confidence intervals and combine all values in matrix.
upper <- slopes + 1.96*se
lower <- slopes - 1.96*se
variable.data <- cbind(variable, slopes, upper, lower)

@@

# EXTENSION

Our extension was inspired by Andreas Wimmer and Yuval Feinstein's 2010 paper "The Rise of the Nation-State Across the World, 1816-2001."  This paper seperately tests the prevailing explanatory theories of the rise of the nation-state as the dominant form of governance. Our extension takes a similar approach, testing the major theories of support for extreme right parties in the Western European context in separate models. We operationalize the variables in these theories using the available measures from the original data set.

In order to obtain a compelling visual representation of the predictive power of different theories and models, we use the method suggested in Brian Greenhill, Michael D. Ward and Audrey Sacks (2011) "The Separation Plot: A New Visual Method for Evaluating the Fit of Models with Categorical Outcomes",  American Journal of Political Science 55 (3): 1-13. We use their R package "separationplot".

# Model 1: Adorno Thesis (Personality Traits and Value Orientations)
The first model is based on Theodor Adorno's 1950 theory that support for extreme right parties (in his case, fascist parties) originates from supporters' personality traits and value orientations, which make individuals more receptive to the appeals of extreme right parties.  We operationalize "personality traits and value orientations" using the individual level characteristics from our data set: gender, age, occupation, marital status, and self-placement on a left-right scale.

```{r}
library(separationplot)
library(Hmisc) 
library(zoom)

model_1 <- glm(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + partner + left_right_scale + country - 1, data = data)
summary(model_1)
actual <- model_1$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_1$fitted.values # Create a vector with log-odd values
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
plot_model1 <- separationplot(pred = pred, actual = actual, type = "rect", heading = "Model1: The Adorno Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
plot_model1
```

@@

#Conclusion
There isn't a big range in the predicted probabilities of the model, therefore the predicted probabilities are not very helpful. Separation plot helps a lot to discern the predictive power of the model. As the events of extreme right vote intension increases as we move to the right (i.e. red lines cluster around the end of the graph), the model does a good job of estimating the likelihood of events. However, as the small arrow in the graph shows, the model predicts that half the number of observations will have the intention to vote for an extreme right party. The number of actual events (red lines) is much lower.

# Model 2: Parsons Thesis (Social Disintegration)
Parsons' 1942 theory of social disintegration posits that anomia, a breakdown of social norms, and attendant feelings of anxiety, anger and isolation brings about a longing for strong leadership and rigid ideologies that extreme right parties can provide.  While we do not have measures for the emotions of participants in the studies whose data we use, in our second model we examine the relationship between extreme right voting and the kinds of sentiments about social change that Parsons refers to: satisfaction with democracy and the negative evaluation of the European Union.

```{r}
model_2 <- glm(ext_vote ~ eu_neg + z_dem_satis + country - 1, data = data)
summary(model_2)

actual <- model_2$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_2$fitted.values # Create a function with log-odds
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
plot_model2 <- separationplot(pred = pred, actual = actual, type = "rect", heading = "Model2: The Parsons Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
plot_model2
```

#Conclusion:
Again, there is little variation in our predictive probabilities with this model. Similar to model 1, most events are aligned with higher predictive probabilities on the right hand of the graph. Yet Model 1 has slightly better predictive power as the very end of the graph is likely to be clustered with event than non-events as in Model 2.

# Model 3: Group Conflict Model (Various theories)
Group conflict theories encompass a variety of theories.  Some emphasize the role of instrumental rationality, arguing that xenophobia results from a contest between immigrants and residents over scarce resources like jobs and welfare benefits. Others emphasize the scapegoating of immigrants as perceived threats to residents' status, power or resources.  In this last model, we use relevant variables representing both the availability of resources like jobs and welfare benefits and the prevalence of immigration: unemployment rates, unemployment benefits, number of asylum seekers, saliance and variance of right-wing issues.

```{r}
model_3 <- glm(ext_vote ~ z_asylumseeker + z_unemp + z_asylumseeker:z_unemp + z_replacement + z_replacement:z_unemp + z_replacement:z_asylumseeker + country - 1, data = data)

summary(model_3)
actual <- model_3$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_3$fitted.values # Create a function with log-odds
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
plot_model3 <- separationplot(pred = pred, actual = actual, type = "rect", heading = "Model3: The Group Conflict Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

#Conclusion:
The majority of events (intention to vote for ER) are clustered at the end right end of the model - which is in line with a good predictive model. However, in this model the distribution of events is very dispersed compared to other two models.

Although there is very little difference in the separation plots for all three models, our conclusion is that model 1 (the Adorno thesis which emphasizes individual-level characteristics and personal attitudes) has the most predictive power.
